The kinematics of the Local Group in a cosmological context

### 2013-02-08 paper/ms.tex
added line breaks

({\bf Jaime - please be more spacific here. How big is your sample? The $10^6$ that appears down the text?. Yehuda- I corrected the whole paragraph. The results here have nothing to do with the montecarlo.})

@@

The difference between the left and right panels of Figure \ref{fig:rt} reflects  the tension found from the full and the sub- sample of the LG-like objects. {\bf Yehuda, What tension??? The tension comes between theory and observation, not between the two theoretical results.}%YH

@@

This formulation has a clear theoretical advantage if one considers the angular momentum and the mechanical energy as quasi-conserved dynamical quantities. This means that, after some formation time, these quantities do not significantly evolve as it is the case for the radial and tangential velocities. 
%YH: the following statement doesn't make much sense. Everything needs to be derived from observations.
%JF: I keep because it is important to mention that the observational uncertainties will be larger.
Conversely, there is an observational disadvantage. These quantities are derived from very different kinds of observations, increasing the uncertainties in their final determination.

A possible conflict between the likelihood of the LG in the $\Lambda$CDM model as viewed from the radial and tangential velocities as opposed to the angular momentum and energy perspectives would point to a particular phase on the LG two-body orbit. Given the masses, velocity and position of the observed MW and M31 (TAble \ref{table:1} ) the reduced energy and angular momentum are calculated. The Monte Carlo sampling is used to to estimate the uncertainty in these quantities. %YH

@@

Inspection of Figure \ref{fig:EJ}  shows that the $e_{\rm tot}$  and $l_{\rm orb}$ constraints are less restrictive compared with the radial and tangential velocity constraints. For the unconstrained full (sub-) sample 15\% (24\%)  of the LG-like pairs obey the 1-$\sigma$ observational constraints. All the constrained LGs of the full 
%JF
%and sub- 
%Not correct, there are 2 constrained pairs that cannot even make it into the sub samples.
samples are consistent with the 1-$sigma$ observational uncertainty.
%JF
%this point is important:
However a reduction by a factor of $2$ in the uncertainty of the tangential velocity showing that the tangential velocity is below $17$\kms would 
bring these pairs and the $\Lambda$CDM expectation outside the 1-$\sigma$ observational uncertainty.

@@

The spin parameter of the simulated LG-like objects is skewed towards higher values compared with the observationally estimated value. 
Inspection of Figure \ref{fig:lambda} shows that the $\lambda$ distribution is close to a lognormal one and thus resembles the distribution of the spin parameter of DM halos in simulation ({\bf Jaime - please provide a reference}) . 
The estimated value of $\log_{10}\lambda$ of the observed LG is $-1.76$, compared with the mean value of 
 $\log_{10}\lambda$ of $-1.51$ for the full sample and $-1.38$ for the subsample. The values of the spin parameter of the 3 constrained LGs are

$\log_{10}(\lambda_{\rm CLUES})=-2.21$, $-1.72$ and $-1.65$, with the latter value belonging to the sole constrained LG of the subsample.

% 
It follows that the observationally estimated energy and angular momentum and the estimated spin parameter impose similar constraints on the simulations and the standard model of cosmology. 
%JF. Yehuda, I don't understaind this sentence. 
This implies that the observed LG does not conspire to have  ($\Lambda$CDM) typical mass, angular momentum and energy and yet have an atypical spin parameter.
%YH

\section{Conclusions}
We have presented a comparison between the observed kinematics for the M31 in the galactocentric restframe and the expectations for a large N-body cosmological simulation in the $\Lambda$CDM cosmology. In the simulation we select a sample of halo pairs in the mass range $5\times 10^{11}<M_{h}/\hMsun<5\times 10^{12}$ that closely match the isolation conditions of the Local Group from other massive structures. While the observations show that M31 moves towards us 
on a highly eccentric orbit,%YH
 the simulation shows that the most common configuration at $z=0$ has values $v_{\rm rad, \Lambda CDM}=-65\pm5$\kms and $v_{\rm tan, \Lambda CDM}=62\pm5$ \kms.

@@

The MultiDark Database used in this paper 
%YH:
%and the web application providing online access to it were 
was %YH
constructed as part of the activities of the German Astrophysical Virtual Observatory as result of a collaboration between the Leibniz-Institute for Astrophysics Potsdam (AIP) and the Spanish MultiDark Consolider Project CSD2009-00064. The Bolshoi simulation was run on the NASA's Pleiades supercomputer at the NASA Ames Research Center.

>>>>

@@

The difference between the left and right panels of Figure
\ref{fig:rt} reflects  the tension found from the full and the sub-
sample of the LG-like objects. {\bf Yehuda, What tension??? The
  tension comes between theory and observation, not between the two
  theoretical results.}

@@

This formulation has a clear theoretical advantage if one considers
the angular momentum and the mechanical energy as quasi-conserved
dynamical quantities. This means that, after some formation time,
these quantities do not significantly evolve as it is the case for the
radial and tangential velocities.

Conversely, there is an observational disadvantage. These quantities
are derived from very different kinds of observations, increasing the
uncertainties in their final determination.

A possible conflict between the likelihood of the LG in the
$\Lambda$CDM model as viewed from the radial and tangential velocities
as opposed to the angular momentum and energy perspectives would point
to a particular phase on the LG two-body orbit. Given the masses,
velocity and position of the observed MW and M31 (TAble \ref{table:1}
) the reduced energy and angular momentum are calculated. The Monte
Carlo sampling is used to to estimate the uncertainty in these
quantities.

@@

Inspection of Figure \ref{fig:EJ}  shows that the $e_{\rm tot}$  and
$l_{\rm orb}$ constraints are less restrictive compared with the
radial and tangential velocity constraints. For the unconstrained full
(sub-) sample 15\% (24\%)  of the LG-like pairs obey the 1-$\sigma$
observational constraints. All the constrained LGs of the full
samples are consistent with the 1-$sigma$ observational uncertainty.

However a reduction by a factor of $2$ in the uncertainty of the tangential velocity showing that the tangential velocity is below $17$\kms would 
bring these pairs and the $\Lambda$CDM expectation outside the
1-$\sigma$ observational uncertainty.

@@

The spin parameter of the simulated LG-like objects is skewed towards higher values compared with the observationally estimated value. 
Inspection of Figure \ref{fig:lambda} shows that the $\lambda$
distribution is close to a lognormal one and thus resembles the
distribution of the spin parameter of DM halos in simulation ({\bf
  Jaime - please provide a reference}) .  
The estimated value of $\log_{10}\lambda$ of the observed LG is
$-1.76$, compared with the mean value of $\log_{10}\lambda$ of $-1.51$
for the full sample and $-1.38$ for the subsample. The values of the
spin parameter of the 3 constrained LGs are $\log_{10}(\lambda_{\rm
  CLUES})=-2.21$, $-1.72$ and $-1.65$, with the latter value belonging
to the sole constrained LG of the subsample.

It follows that the observationally estimated energy and angular momentum and the estimated spin parameter impose similar constraints on the simulations and the standard model of cosmology. 
%JF. Yehuda, I don't understaind this sentence. 
%This implies that the observed LG does not conspire to have  ($\Lambda$CDM) typical mass, angular momentum and energy and yet have an atypical spin parameter.
%YH

\section{Conclusions}
We have presented a comparison between the observed kinematics for the
M31 in the galactocentric restframe and the expectations for a large
N-body cosmological simulation in the $\Lambda$CDM cosmology. In the
simulation we select a sample of halo pairs in the mass range $5\times
10^{11}<M_{h}/\hMsun<5\times 10^{12}$ that closely match the isolation
conditions of the Local Group from other massive structures. While the
observations show that M31 moves towards us  on a highly eccentric orbit,
 the simulation shows that the most common configuration at $z=0$ has
 values $v_{\rm rad, \Lambda CDM}=-65\pm5$\kms and $v_{\rm tan,
   \Lambda CDM}=62\pm5$ \kms.

@@

The MultiDark Database used in this paper  was constructed as part of
the activities of the German Astrophysical Virtual Observatory as
result of a collaboration between the Leibniz-Institute for
Astrophysics Potsdam (AIP) and the Spanish MultiDark Consolider
Project CSD2009-00064. The Bolshoi simulation was run on the NASA's
Pleiades supercomputer at the NASA Ames Research Center.
